---
title: "Model formulations"
author: "Nicholas Reich"
date: "February 21, 2017"
output: pdf_document
---

```{r, eval=FALSE}
devtools::install_github('cdcepi/tsir-sims')
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSIRsim)
```

## Some notation

Let $Y_t$ be an observed case count at time $t$ from a latent, unobserved process $Z_t$. Also, let the population be $N_t$. 

Several processes govern how $Z$ is observed. Specifically, we assume that $\pi^{T|Z}$ is the probability of getting tested given that you have disease and $Se$ is the sensitivity of the test (i.e. the probability that the test will be positive given that you have the disease). Also, if $\pi^Z_t$ is the probability at time $t$ of having the disease, then a simple observation model is:
$$Y_t \sim Binom(\pi^Z_t \cdot \pi^{T|Z} \cdot Se, N_t)$$

We can also use a simple model for the disease process. E.g. 
$$ Z_t \sim NB(\frac{r}{r+\lambda_t}, r) $$
where the mean of the negative binomial is then $\lambda_t$.
Also, 
$$ \lambda_t = \beta \cdot Z_{t-1} \cdot S_t / N_t $$
and
$$ S_t = S_{t-1} - Z_{t-1}. $$
Further, $Z_0$ is an estimated parameter, with $S_1 := N_t-Z_0$.

## A simple simulation

```{r}
tsir_sim <- function(beta, r, N, nsteps=104, plot=TRUE) {
    Z_0 <- dpois(1, round(N/10))
    Z <- S <- p <- lam <- rep(NA, nsteps)
    S[1] <- N - Z_0
    lam[1] <- beta * Z_0 * S[1] / N
    p[1] <- r/(r+lam[1]) 
    Z[1] <- rnbinom(1, size=p[1], prob=r)
    for(t in 2:nsteps){
        S[t] <- S[t-1] - Z[t-1]
        lam[t] <- beta * Z[t-1] * S[t] / N
        p[t] <- r/(r+lam[t]) 
        Z[t] <- rnbinom(1, size=p[t], prob=r)
    }
    plot(Z)
    return(Z)
}
tsir_sim(beta=2, r=0.5, N=10000, nsteps=104)
```

## Open questions

 - If we simulate from a real SIR model, does this simple model fit the data ok? What essential features of the data does this capture?
 - Is the fitting dependent on a system starting from a relatively naive state, with little prevailing immunity? Seems like that state is implicit in the `S[1] <- N - Z_0` statement.
 - We could try simulating and estimating datasets where we vary the following features to see how they impact estimation: (1) $S_0/N$   , (2) $N$, (3) $\mathbb E \left [ \sum_t Z_t \right] $.
 - Add a version of sim_tsir with noise

## A better simple simulation
```{r}
sample.z = sim_tsir_simple(pop = 3.5e6, 
                           r.init = 0, 
                           z.init = 10, 
                           beta = 1.5, 
                           p.obs = 0.2, 
                           ntimes=52)
plot(sample.z$obs)
sum(sample.z$obs)/sample.z$pop
```

```{r}

test <- fit_sim_data(pop = 0.5e6, n.sim=1000, 
                     times=create.time.ref())

```


